---
title: "Anti-psychotic drug analysis"
author: "Dr Greig Russell"
date: "2 March 2018"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

#Introduction

This analysis describes the anti-psychotic usage of 100 patients from the mental health service. It will describe the characteristics of the sample and the anti-psychotic medication prescribed for them. This description will provide the base to understand the relationships between anti-psychotic prescriptions and features of the population.

#Summary
This study has found that more females are admitted with psychosis, and these females are statistically significantly older (p = 0.008). 22% of this sample received greater than the recommended dose of anti-psychotics whilst an inpatient. The more varieties of anti-psychotic medication that a patient was prescribed, the more likely they were to exceed the maximum recommended dose (p = 0.002). There was no significant relationship between total dosage and gender (p = 0.07) or ethnicity (p = 0.22). In the sub-group who were prescribed supra-maximal dosage there was no difference by Gender (p = 0.49) or Ethnicity (p = 0.53)

There is a relationship between total dosage prescribed as an in-patient and dosage on discharge (p = 0.01). NZ Maori are statistically more likely to receive intra-muscular medication than NZ European (p = 0.04). 


```{r load_the_libraries, warning=FALSE, message=FALSE}
# Load the tidyverse
library(tidyverse)

# Load any axillary libraries
library(nullabor)
library(car)
library(gmodels)
library(broom)
```

```{r load_the_data, warning=FALSE, message=FALSE}
Drug <- read.csv("/home/greig/R-projects/Clinical_audits/Drug/Anti-psychotic/Drug.csv", na.strings = "")

Drug <- Drug %>% 
  rename(Drug1_NZFpercentage = Drug1_NZFpercentage.)
```

```{r basic_demographic_data, warning=FALSE, message=FALSE}
# summary(Drug)
```

\pagebreak

#Sample demographics

##Age distribution
```{r age1, warning=FALSE, message=FALSE}
ggplot(data = Drug) +
  geom_density(aes(x = Age), fill = "red") +
  labs(title = "Density plot of cohort age") +
  theme_minimal()
```

```{r Age_model, warning=FALSE, message=FALSE}
Boy <- Drug %>%
  filter(Gender == "Male")

Girl <- Drug %>%
  filter(Gender == "Female")

Count_boy <- Boy %>%
  summarise(Count = n(), Mean_age = mean(Age, na.rm = TRUE))

Count_girl <- Girl %>%
  summarise(Count = n(), Mean_age = mean(Age, na.rm = TRUE))

Age_model <- tidy(t.test(Boy$Age, Girl$Age))

```

As the Box plot illustrates there are more females (n = ```r Count_girl$Count```) with a higher mean age of ```r round(Count_girl$Mean_age, digit = 0)``` in this sample than males (n = ```r Count_boy$Count```) whose mean age is ```r round(Count_boy$Mean_age, digit = 0)```. This age difference is significant (p = ```r round(Age_model$p.value, digits = 3)```).  

##Gender
```{r gender1, warning=FALSE, message=FALSE}
Drug %>% 
  group_by(Gender) %>% 
  summarise(Count = n()) %>% 
  ggplot(aes(x = Gender, y = Count, fill = Gender)) +
  geom_col() +
  labs(title = "Gender distribution") +
  theme_minimal()
```

##Age by Gender
```{r gender2, warning=FALSE, message=FALSE}
ggplot(data = Drug) +
  geom_boxplot(aes(x = Gender, y = Age, fill = Gender)) +
  labs(title = "Boxplot of cohort age by gender") +
  theme_minimal()
```

\pagebreak

#Anti-psychotic presecribing patterns

##Total regular anti-psychotic dosage by number of drug types
```{r Drug1, warning=FALSE, message=FALSE}
Poly <- Drug %>% 
  filter(Poly_Num_drugs < 4)
Poly$Poly_Num_drugs <- as.factor(Poly$Poly_Num_drugs)

Excess <- Drug %>%
  filter(Poly_sum_reg_NZFpercentage > 1) %>% 
  summarise(Count = n())
```

In this cohort of `r nrow(Drug)` clients, `r Excess$Count` received in excess of the NZULM recommended maximum dosage of regular anti-psychotic medications not including "prn" or as required supplementary anti-psychotic medication dosages.

```{r Drug2, warning=FALSE, message=FALSE}
ggplot(data = Poly) +
  geom_boxplot(aes(x = Poly_Num_drugs, y = Poly_sum_reg_NZFpercentage, fill = Poly_Num_drugs)) +
  labs(title = "Total dosage of presecribed anti-psychotic by number of drug types used") +
  theme_minimal()
```


Prior to testing for significance by ANOVA, Levene test performed to conform homogeneity of variance
```{r levene1, warning=FALSE, message=FALSE}
#Levene test

leveneTest(Poly$Poly_sum_reg_NZFpercentage, Poly$Poly_Num_drugs, center = mean)
```


```{r anova1, warning=FALSE, message=FALSE}
Model1 <- aov(Poly_sum_reg_NZFpercentage ~ Poly_Num_drugs, data = Poly)
summary(Model1)
```

Supporting the hypothesis that the number of drug types prescribed is significantly positively correlated with the total dosage of anti-psychotics received per day.


##Total regular anti-psychotic dosage by gender
```{r gender1a, warning=FALSE, message=FALSE}
ggplot(data = Poly) +
  geom_boxplot(aes(x = Gender, y = Poly_sum_reg_NZFpercentage, fill = Gender)) +
  labs(title = "Total dosage of presecribed anti-psychotic by gender of the client") +
  theme_minimal()
```

Again testing for the homogeneity of the variance;
```{r gender2a, message=FALSE, warning=FALSE}
# Levene test

leveneTest(Poly$Poly_sum_reg_NZFpercentage, Poly$Gender, center = mean)

```

```{r gender3, message=FALSE, warning=FALSE}
Model2 <- aov(Poly_sum_reg_NZFpercentage ~ Gender, data = Poly)
summary(Model2)
```

##Relationship between total anti-psychotic drug usage during an admission and ethnicity
```{r drug_ethnicity, warning=FALSE, message=FALSE}
Drug %>% 
  ggplot(aes(x = Ethnicity, y = Poly_sum_reg_NZFpercentage, fill = Ethnicity)) +
  geom_boxplot() +
  labs(title = "Total antipsychotic drug usage by ethnicity", y = "Total inpatient anti-psychotic dose as % NZULM") + 
  theme_minimal()
```

```{r drug_ethinicity2, warning=FALSE, message=FALSE}
leveneTest(Poly$Poly_sum_reg_NZFpercentage, Poly$Ethnicity, center = mean)
```

```{r drug_ethnicity3, warning=FALSE, message=FALSE}
model4 <- aov(Poly_sum_reg_NZFpercentage ~ Ethnicity, data = Drug)
summary(model4)
```

Suggesting there is no difference between the drug dosage received during an admission on the basis of ethnicity

\pagebreak

##Analysis of the super maximal dosage cohort

```{r Super_max, warning=FALSE, message=FALSE}
Super_max_Maori <- Drug %>%
  filter(Poly_sum_reg_NZFpercentage > 1,
         Ethnicity == "NZ Maori")

Super_max_European <- Drug %>%
  filter(Poly_sum_reg_NZFpercentage > 1,
         Ethnicity == "NZ European")

Super_max_boy <- Drug %>%
  filter(Poly_sum_reg_NZFpercentage > 1,
         Gender == "Male")

Super_max_girl <- Drug %>%
  filter(Poly_sum_reg_NZFpercentage > 1,
         Gender == "Female")

Drug %>%
  filter(Poly_sum_reg_NZFpercentage > 1) %>% 
  count(Ethnicity)

Drug %>%
  filter(Poly_sum_reg_NZFpercentage > 1) %>%
  count(Gender)

t.test(Super_max_Maori$Poly_sum_reg_NZFpercentage, Super_max_European$Poly_sum_reg_NZFpercentage)

t.test(Super_max_girl$Poly_sum_reg_NZFpercentage, Super_max_boy$Poly_sum_reg_NZFpercentage)
```



##Relationship between drug levels during inpatient and on discharge
```{r discharge1, warning=FALSE, message=FALSE}
ggplot(data = Poly, aes(x = Poly_sum_reg_NZFpercentage, y = Discharge_Reg_NZF_Percentage)) +
  geom_point() +
  geom_smooth() +
  labs(Title = "Relationship between inpatient drug levels and those following discharge")
```

```{r discharge2, warning=FALSE, message=FALSE}
Model3 <- lm(Discharge_Reg_NZF_Percentage ~ Poly_sum_reg_NZFpercentage, data = Poly)
summary(Model3)
```

Revealing a highly significant relationship between the doses of drugs received during an admission and those received following discharge

##Relationship between Ethnicity and Depot anti-psychotic drug usage
```{r depot, warning=FALSE, message=FALSE}
Poly <- Drug %>%
  filter(Poly_Num_drugs < 4, (Ethnicity == "NZ European" | Ethnicity == "NZ Maori"))
Poly$Drug1_route <- str_to_upper(Poly$Drug1_route)
CrossTable(Poly$Ethnicity, Poly$Drug1_route, prop.r = FALSE, prop.c = FALSE, prop.chisq = FALSE, chisq = TRUE)
```

Revealing that NZ Maori are statistically more likely to receive depot anti-psychotic medication.

