---
title: "Anti-psychotic drug analysis"
author: "Dr Greig Russell"
date: "2 March 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

#Anti-psychotic drug analysis

This analysis describes the anti-psychotic usage of 100 patients from the mental health service. It will describe the characteristics of the sample and the anti-psychotic medication presecribed for them. This description will provide the base to understand the relationships between anti-psychotic presecriptions and features of the population.

##Summary

```{r load_the_libraries, warning=FALSE, message=FALSE}
# Load the tidyverse
library(tidyverse)

# Load any axillary libraries
library(nullabor)
library(car)
```

```{r load_the_data, warning=FALSE, message=FALSE}
Drug <- read.csv("/home/greig/R-projects/Clinical_audits/Drug/Anti-psychotic/Drug.csv", na.strings = "")

Drug <- Drug %>% 
  rename(Drug1_NZFpercentage = Drug1_NZFpercentage.)
```

```{r basic_demographic_data, warning=FALSE, message=FALSE}
summary(Drug)
```


##Sample demographics

###Age distribution
```{r age, warning=FALSE, message=FALSE}
ggplot(data = Drug) +
  geom_density(aes(x = Age), fill = "red") +
  labs(title = "Density plot of cohort age") +
  theme_minimal()
```


###Gender
```{r}
ggplot(data = Drug) +
  geom_boxplot(aes(x = Gender, y = Age, fill = Gender)) +
  labs(title = "Boxplot of cohort age by gender") +
  theme_minimal()
```

